---
title: "week3_exercise"
author: "Aiyana Signer"
date: "2023-05-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("readr") 
library(ggplot2)
library(tidyverse)
library("sf") 
library(SimilarityMeasures)
```

## Prepare Data
```{r}
# import data
posmo <- read_delim("data/posmo_2023-01-01T00_00_00+01_00-2023-05-01T23_59_59+02_00.csv", ",")

# Keep only the necessary columns
posmo <- select(posmo, datetime, lon_x, lat_y)

posmo <- st_as_sf(posmo, coords = c("lon_x","lat_y"), crs = 4326, remove = FALSE, na.fail = FALSE) |>
  st_transform(2056)

head(posmo)
```

```{r}
# coordinates in separate columns
posmo_coordinates <- st_coordinates(posmo)

posmo <- cbind(posmo, posmo_coordinates)
```

```{r}
posmo_filter <- posmo |>
    filter(as.Date(datetime) == "2023-04-14")
```

## Task 1: Segmentation
```{r}
#b. Measure the distance from every point to every other point
posmo_filter <- posmo_filter |>
  mutate(
    nPlus1 = sqrt((lead(X,1)-X)^2 + (lead(Y,1)-Y)^2),
    nPlus2 = sqrt((lead(X,2)-X)^2 + (lead(Y,2)-Y)^2),
    nMinus1 = sqrt((lag(X,1)-X)^2 + (lag(Y,1)-Y)^2),
    nMinus2 = sqrt((lag(X,2)-X)^2 + (lag(Y,2)-Y)^2)
    )

# row-wise step mean
posmo_filter <- posmo_filter |> 
  rowwise() |> 
  mutate( stepMean = mean(c(nMinus2, nMinus1, nPlus1, nPlus2)) 
          ) |> 
  ungroup()

#c. Remove “static points”
# These are points where the average distance is less than a given threshold. This segments the trajectory into subtrajectories
posmo_filter <- posmo_filter |> 
  ungroup() |> 
  mutate(static = stepMean < mean(stepMean, na.rm = TRUE))

posmo_filter_stat <- posmo_filter |>
  filter(!static)

posmo_filter_stat |> 
  ggplot(aes(X, Y)) + 
  geom_path() + 
  geom_point() + 
  coord_fixed() + 
  theme(legend.position = "bottom")
```

## Task 2: Specify and apply threshold d
```{r}
# summary
summary(posmo_filter$stepMean)

# histogram of steplength
ggplot(posmo_filter, aes(stepMean))+
  geom_histogram(binwidth = 20) + 
  labs(x = "Steplength [m]",
       y = "Count") +
  geom_vline(xintercept = mean(posmo_filter$stepMean, na.rm = TRUE)) +
  theme_minimal()

# apply threshold d
posmo_filter <- posmo_filter %>%
    ungroup() %>%
    mutate(static = stepMean < mean(stepMean, na.rm = TRUE))
```

## Task 3: Visualize segmented trajectories
```{r}
ggplot(posmo_filter, aes(X, Y, color = static)) + 
  geom_point()+
  geom_path() + 
  coord_equal()
```

## Task 4: Segment-based analysis

```{r}
# define unique ID function
rle_id <- function(vec) {
    x <- rle(vec)$lengths
    as.factor(rep(seq_along(x), times = x))
}

# apply rle_id function
posmo_filter <- posmo_filter |>
    mutate(segment_id = rle_id(static))

# only keep certain columns
posmo_filter <- select(posmo_filter, datetime, X, Y, geometry, static, segment_id)

head(posmo_filter)
```

```{r}
# filter out static trajectories
posmo_filter_nonstatic <- posmo_filter %>%
    filter(!static)
posmo_filter_nonstatic %>%
  ggplot(aes(X, Y, color = segment_id)) +
  geom_path() +
  geom_point() +
  coord_equal()+
  theme_minimal()+
  theme(legend.position = "none")+ # do not show legend
  ggtitle("All segments (uncleaned)")

# remove short segments
# calculate time difference between two fixes
posmo_filter_nonstatic <- posmo_filter_nonstatic |> 
  mutate(
    timediff_s = as.numeric(lead(datetime, 1)-datetime)
  )

# calculate total time difference per segment
posmo_filter_nonstatic <- posmo_filter_nonstatic |> 
  group_by(segment_id) |> 
  mutate(
    segmenttime_s = sum(timediff_s)
  ) |> 
  ungroup()

# set time threshold
time_threshold <- 60*5 #convert 5 minutes to seconds

posmo_filter_nonstatic_longseg <- posmo_filter_nonstatic |> 
  filter(!segmenttime_s < time_threshold) # removes rows where total segment time less than 5 minutes
```

```{r}
# visualize without short segments (duration less than 5 minutes)
posmo_filter_nonstatic_longseg %>%
  ggplot(aes(X, Y, color = segment_id)) +
  geom_path() +
  geom_point() +
  coord_equal()+
  theme_minimal()+
  ggtitle("Long segments (removed segments < 5 minutes)")
```

## Task 5: Similarity measures
```{r}
# import data
ped <- read_delim("data/pedestrian.csv", ",")

head(ped)
```

```{r}
# visualize pedestrian data
ggplot(ped, aes(x = E, y = N, color = factor(TrajID))) +
  geom_point() + 
  geom_path() +
  labs(title = "Visual comparison of the 6 trajectories", x = "E", y = "N", color = "Trajectory ID") +
  facet_wrap(~ TrajID) # create one subplot per trajectory +
  coord_equal()
  
```


## Task 6: Calculate similarity

Before visualizing your results think about the following: Which two trajectories to you perceive to be most similar, which are most dissimilar?

Based on the shape, trajectories 1,2, 3 and 6 seem rather similar, while Trajectory 4 and 5 stand out as being different. If one also takes into account the spacing between the points, then Trajectory 1 and 6 are quite similar.

```{r}
# create a list of data frames for each trajectory
traj_list <- split(ped, ped$TrajID)

# convert each data frame to a matrix
traj_mat_list <- lapply(traj_list, function(x) as.matrix(x[, c("E", "N")]))

# compute the similarity metric between trajectory 1 and trajectories 2 to 6
similarity_DTW <- sapply(traj_mat_list[-1], function(traj) DTW(traj_mat_list[[1]], traj))
similarity_EditDist <- sapply(traj_mat_list[-1], function(traj) EditDist(traj_mat_list[[1]], traj))
similarity_Frechet  <- sapply(traj_mat_list[-1], function(traj) Frechet(traj_mat_list[[1]], traj))
similarity_LCSS <- sapply(traj_mat_list[-1], function(traj) {LCSS(traj_mat_list[[1]], traj, pointSpacing = 100, pointDistance = 100, errorMarg = 0.5)})

# turn results into data frames
df_DTW <- data.frame(Trajectory = names(similarity_DTW), Similarity = similarity_DTW, Measure = "DTW")
df_EditDist <- data.frame(Trajectory = names(similarity_EditDist), Similarity = similarity_EditDist, Measure = "EditDist")
df_Frechet <- data.frame(Trajectory = names(similarity_Frechet), Similarity = similarity_Frechet, Measure = "Frechet")
df_LCSS <- data.frame(Trajectory = names(similarity_LCSS), Similarity = similarity_LCSS, Measure = "LCSS")

# merge data frames
df_similarities <- bind_rows(df_DTW, df_EditDist, df_Frechet, df_LCSS)

# plot
ggplot(df_similarities, aes(x = Trajectory, y = Similarity, fill = Trajectory)) +
  geom_bar(stat = "identity") + 
  facet_wrap(~ Measure, scales = "free") +
  labs(x = "Trajectory", y = "Similarity", title = "Computed Similarities using different measures between Trajectory 1 to all other Trajectories") +
  theme(legend.position = "none")
```


